---
title: "Employee Survey Analysis"
output: html_notebook
---

For our Employee Survey Analysis we'll break the project into three steps
 
 Step 1: EDA
  - loading our data into our working environment
  - load necessary packages for project
  - understanding our data (dimensions/rows & columns/data types etc.)
  
  
  Step 2: Data Transformation
  - Should know which transformations are needed based on above step
  - Applying data transformation steps into new dataframe
  - Checking results for desired outcome
  
  Steps 3: Data Analysis
  - Estimating internal consistency reliability using Cronbach's Alpha (ensuring there's intercorrelation among results by theme)
  - Creating a composite variable based on a multi-item measure (aggregating scores by theme)
 

```{r }
# packages used

library(dplyr)
library(tidyverse)
```


```{r }
# data used

## first survey responses
EmpSurvData<-read.csv("https://raw.githubusercontent.com/davidcaughlin/R-Tutorial-Data-Files/master/EmployeeSurveyData.csv")


```



```{r }
# initial overview of data

head(EmpSurvData)
summary(EmpSurvData)


# looks like we have 
  ## dept
  ## emp ID
  ## emp mgr id
  ## four categories of the questions (turnover intention/job satisfaction/engagement/exposure to incivility items)
```

Time to aggregate & segment our data, to understand our outcomes better


Response count by dept sort by response count

```{r }

EmpSurvData %>% 
  group_by(Dept. = Unit) %>% 
  summarize(ResponseCount = n()) %>% 
  ungroup() %>%
  arrange(ResponseCount)
 

```
 Response count by dept & mgr
```{r}

EmpSurvData %>% 
  group_by("Dept." = Unit, "Dept. Mgr" = Supervisor) %>% 
  summarize(ResponseCount = n()) %>% 
  ungroup()
```

Let's add two aggregate columns to a new dataframe

```{r}

agg_empsurvdata<- EmpSurvData %>% 
  group_by(Unit) %>% 
  mutate("Response Count" = n()) %>% 
  ungroup()

head(agg_empsurvdata)
```

