---
title: "Employee Survey Analysis"
output: html_notebook
---

For our Employee Survey Analysis we'll break the project into a couple of parts
 
 Part 1: Import & Inspect our data
  - loading our data into our working environment
  - load necessary packages for project
  - check quality of data (nulls/errors/duplicates etc.)
  
  
 Part 2: Clean & Transform our data
  - Rename/remove columns & filter any data not needed
  - Create/Recode columns needed for analysis
  - Pivot columns in order to pass into BI tools
  - BONUS: Data Imputation using MICE
  
 Part 3: Data Analysis
  - Estimating internal consistency reliability using Cronbach's Alpha (ensuring there's intercorrelation among results by theme)
  - Creating a composite variable based on a multi-item measure (aggregating scores by theme)
 
====================================================================================================

Part 1: Import & Inspect 

```{r }

## loading data
survey_raw<-read.csv("https://raw.githubusercontent.com/dlopez972/Employee_Survey_Analysis/main/sample_employeesurvey_table.csv")

surv_key_raw<-read.csv("https://raw.githubusercontent.com/dlopez972/Employee_Survey_Analysis/main/sample_employeesurvey_key.csv")

## inspecting data
head(survey_raw)
summary(survey_raw)
length(unique(survey_raw$EID))

                        ## 199 records w/ 20 columns
                        ## Issue 1: First & Last Name all null values (drop columns)
                        ## Issues 2: See null values for 43 records in their responses 
                        ## checked for duplicate EIDs, none to be found

## packages used

library(dplyr)
library(tidyverse)
library(mice)
library(psych)
```


====================================================================================================

Part 2: Clean & Transform
 
Steps for transformations: 
 step 1: pivot question columns
 step 2: replace values for certain columns
 step 3: create new column "Responded?"
 step 4: remove unnecessary columns
 step 5: rename remaining columns
 Bonus Step: Data imputation 


```{r}
# let's first create a copy of our data, and keep the raw
survey_copy<-survey_raw

# now we can start our steps

survey_transformed <- survey_copy %>% 
  select(-First.Name, -Last.Name) %>% 
  pivot_longer(starts_with("Q"), names_to = "Questions", values_to = "Response") %>% 
  mutate(
    "Performance Score" = Performance, 
    Gender = case_when(Gender == 1 ~ "Female",
                       Gender == 0 ~ "Male",
                       TRUE ~ "Unknown"),
    Ethnicity = case_when(Ethnicity == 0 ~ "Black",
                          Ethnicity == 1 ~ "Asian",
                          Ethnicity == 2 ~ "White",
                          Ethnicity == 3 ~ "Hispanic",
                          Ethnicity == 4 ~ "Multi-Racial",
                          TRUE ~ "Unknown" ),
    Mgmt.Level = case_when(Mgmt.Level == 1 ~ "Jr. Associate",
                           Mgmt.Level == 2 ~ "Associate",
                           Mgmt.Level == 3 ~ "Sr. Associate",
                           Mgmt.Level == 4 ~ "Principal Associate",
                           Mgmt.Level == 5 ~ "Manager",
                           Mgmt.Level == 6 ~ "Sr. Manager",
                           Mgmt.Level == 7 ~ "Director",
                           Mgmt.Level == 8 ~ "Sr. Director",
                           Mgmt.Level == 9 ~ "VP"),
    Performance  = case_when(Performance  == 1 ~ "Needs Improvement",
                            Performance == 2 ~ "Below Expectations",
                            Performance == 3 ~ "Meeting Expectations",
                            Performance == 4 ~ "Exceeding Expectations",
                            Performance == 5 ~ "Excellent"), 
    "Responded?" = ifelse(is.na(Response),0,1)
    )

head(survey_transformed)
survey_transformed
summary(survey_transformed)


-----------------------------
  
## Bonus: Data Imputation

md.pattern(survey_copy)

survey_imputed <- mice(survey_copy, m=5,method = "rf")
summary(survey_imputed)
survey_imputed$imp$Q5


```






====================================================================================================

Part 3: Calculate Internal Consistency Reliability using Cronbach's Alpha
```{r}
view(EmpSurvKey)


jobsat<-alpha(EmpSurvData[,c("Q1","Q2","Q3")])
head(jobsat)

turnover<-alpha(EmpSurvData[,c("Q4","Q5","Q6")])
head(turnover)

engagement<-alpha(EmpSurvData[,c("Q7","Q8","Q9","Q10")])
head(engagement)

```

